Field-informed Reinforcement Learning for Learning Large-Scale Collective Tasks

### 2023-05-16 paper.tex
more comments and details on scenarios

 $f_v = (\theta^v_t, o^v_t)$. \lukas{we used $o_v$ to describe the observation of node $v$ in section II.C - so this would not include field information, right?}
 \lukas{we have not introduced $G'_t$ yet}

@@

 \lukas{Global information is not available to the agent, right? The agent only provide locally aggregated information (resulting from the AC part)?}

@@

\subsection{Scenario}
Figure~\ref{fig:example} illustrates the scenario used during the learning, 
which consists of a set of 25 agents and a single, stationary phenomenon with a 3D Gaussian 
Each agent can perceive the presence
 of the phenomenon of interest through an installed sensor $\zeta_v$ with $v \in \mathbb{V}$. 
 (e.g., camera, temperature sensor, etc.). 
%
Additionally, each agent has a coverage range $\omega$ (fixed to 7.5 meters) 
 that describes the area it can monitor. 
 Each agent can only communicate directly with its own neighbourhood $\mathcal{N}$, 
 which in this case depends on a $\mathbb{O}$ range fixed to 30 meters. %\lukas{$\mathbb{O}$  not defined.}
From the neighbourhood, 
 each agent can also perceive the direction of the other agents.
%
Each agent moves following a certain action composed of two components ($\alpha$, $\theta$) 
 which respectively describe the angle and intensity of the movement (i.e., the velocity vector).
 %\lukas{drone, robot, agent, node...}
\subsection{Goal}
 The objective of this scenario is threefold:
\begin{enumerate}
\item Maximise the number of agents within the phenomena
\item Minimise the number of agents without a neighbourhood
\item Maximise the coverage of the system
\end{enumerate}

@@

Finally, to maximise the coverage, 
  we define a reward function that favours the maximum distance between the agents
  equal to the coverage range $\omega$.
%
This means that the average distance will tend towards the one expressed by the viewing range of each node:
% 
\begin{equation*}
R^{C}_{v} = 1 - \frac{d_{min}}{\omega}
\end{equation*}
where $d_{min}$ describes and minimum distance of the agent to its neighbourhood. 
%
%\lukas{should this be $1-R^C_v$? or do we want to favour the highest average distance?}
The final reward function for an agent $v$ is defined as:
\begin{equation*}
R_{v} = \frac{R^a_{v} + R^{\mathcal{N}}_{v} + R^{C}_{v}}{3}
\end{equation*}
Specifically, we decided to express the signal as a \emph{regret}
  as it is a more general measure of the quality of the action taken by the agent.
It is simplify evaluated as:
\begin{equation*}
R^e_{v} = R_v - 1
\end{equation*}
%

@@

\subsection{Discussion and Results}
\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/start.png}
      \caption{Initial configuration}
      \label{fig:initial}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/after.png}
      \caption{After 50 steps}
      \label{fig:after}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/end.png}
      \caption{After 200 steps}
      \label{fig:end}
  \end{subfigure}
  \caption{Test results of the proposed approach. 
  The agents go towards the phenomenon and try to cover it without collapsing into a single point.}
  \label{fig:simulations}
\end{figure}
\begin{figure}
  \centering
  \begin{subfigure}[b]{0.49\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-in-time}
    \caption{Coverage in time}
    \label{fig:coverage}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\linewidth}
    \includegraphics[width=\linewidth]{imgs/reward-in-time}
    \caption{Reward in time}
    \label{fig:reward}
  \end{subfigure}
  \caption{Coverage and reward in time}
  \label{fig:coverage-reward}
\end{figure}
%% Create a subfigure of coverage-test, inside-test, inside-two-test and coverage-two-test

\begin{figure*}
  \centering
  \begin{subfigure}[b]{0.49\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-test.pdf}
    \caption{Coverage test}
    \label{fig:coverage-test}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\linewidth}
    \includegraphics[width=\linewidth]{imgs/inside-test}
    \caption{Inside test}
    \label{fig:inside-test}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\linewidth}
    \includegraphics[width=\linewidth]{imgs/inside-two-test}
    \caption{Inside two test}
    \label{fig:inside-two-test}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-two-test}
    \caption{Coverage two test}
    \label{fig:coverage-two-test}
  \end{subfigure}
  \caption{Test results, \ga{todo}}
  \label{fig:test}
\end{figure*}
\section{Conclusion and Future Work}
\ga{Page budget: 0.5}
\label{sec:conclusion}

>>>>

 $f_v = (\theta^v_t, o^v_t)$.
%  \lukas{we used $o_v$ to describe the observation of node $v$ in section II.C - so this would not include field information, right?}
% \lukas{we have not introduced $G'_t$ yet}

@@

Importantly, each agent only has local information from itself and its neighbourhood to utilise in the \ac{GNN}.
% \lukas{Global information is not available to the agent, right? The agent only provide locally aggregated information (resulting from the AC part)?}

@@

\subsection{Scenario}
Figure~\ref{fig:example} illustrates the scenario used during the learning \lukas{is this training the \ac{GNN}? }, 
which consists of a set of 25 agents and a single, stationary phenomenon with a 3D Gaussian 
Each agent can perceive the presence
 of the phenomenon of interest through an installed sensor $\zeta_v$ with $v \in \mathbb{V}$. 
 (e.g., camera, temperature sensor, etc.). 
%
Additionally, each agent has a coverage range $\omega$ (fixed to 7.5 meters) 
 that describes the area it can monitor. 
 Each agent can only communicate directly with its own neighbourhood $\mathcal{N}$, 
 which in this case depends on a $\mathbb{O}$ range fixed to 30 meters. %\lukas{$\mathbb{O}$  not defined.}
From the neighbourhood, 
 each agent can also perceive the direction of the other agents.
%
Each agent moves following a certain action composed of two components ($\alpha$, $\theta$) 
 which respectively describe the angle and intensity of the movement (i.e., the velocity vector). 
 \lukas{$\theta$ used for target weights...} 
 \lukas{not sure if this paragraph should be its own subsection. or at least it needs  different title.}
 For the evaluation, we explore three different types of scenarios. The first type considers only a single, stationary phenomenon. The second type considers multiple phenomena, requiring the agents to make a decision which phenomena to cover. Finally, the third type considers a single moving phenomena which the agents have to follow throughout the simulation. In all three types of scenarios, the shape and underlying distribution is initially unknown to the agents but remains constant for the entire duration of the simulation \lukas{correct?}. We generated random scenarios for each type of scenario. Examples of such types are shown in Figure~\ref{??} \lukas{can we create a figure with at least 3 sample scenarios (1 stationary, 2 split, 1 moving (with a line indicating its trajectory))?} We compare our field-informed reinforcement learning approach against a baseline approach utilising a \ac{MLP} as well as an approach only relying on \acp{GNN}. In all approaches, the underlying neural network (i.e., the \ac{MLP} and the \ac{GNN}) are trained with \lukas{xyz} samples of a single, stationary phenomena. \lukas{we should describe the training a bit.}
 \lukas{do we need more information on the MLP?}

 %\lukas{drone, robot, agent, node...}
\subsection{Goal}
 The objective of this scenario is threefold:
\begin{enumerate}
\item Maximise the number of agents within the phenomena
\item Minimise the number of agents without a neighbourhood
\item Maximise the coverage of the system
\end{enumerate}

@@

Finally, to maximise the coverage, 
  we define a reward function that favours the maximum distance between the agents
  equal to the coverage range $\omega$. This will minimise multiple agents covering a common area.
%
This means that the average distance will tend towards the one expressed by the viewing range of each node:
% 
\begin{equation*}
R^{C}_{v} = 1 - \frac{d_{min}}{\omega}
\end{equation*}
where $d_{min}$ describes and minimum distance of the agent to its neighbourhood. 
%
%\lukas{should this be $1-R^C_v$? or do we want to favour the highest average distance?}
The final reward function for an agent $v$ is defined as:
\begin{equation*}
R_{v} = \frac{R^a_{v} + R^{\mathcal{N}}_{v} + R^{C}_{v}}{3}
\end{equation*}
Specifically, we decided to express the signal as a \emph{regret}
  as it is a more general measure of the quality of the action taken by the agent.
It is simplify evaluated as:
\begin{equation*}
R^e_{v} = R_v - 1
\end{equation*}
%

@@

\subsection{Discussion and Results}
\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/start.png}
      \caption{Initial configuration}
      \label{fig:initial}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/after.png}
      \caption{After 50 steps}
      \label{fig:after}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/end.png}
      \caption{After 200 steps}
      \label{fig:end}
  \end{subfigure}
  \caption{Test results of the proposed approach. 
  The agents go towards the phenomenon and try to cover it without collapsing into a single point.}
  \label{fig:simulations}
\end{figure}
\begin{figure}
  \centering
  \hfill
  \begin{subfigure}[b]{0.96\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-in-time}
    \caption{Coverage in time}
    \label{fig:coverage}
  \end{subfigure}\\
  \hfill
  \begin{subfigure}[b]{0.99\linewidth}
    \includegraphics[width=\linewidth]{imgs/reward-in-time}
    \caption{Reward in time}
    \label{fig:reward}
  \end{subfigure}
  \caption{Coverage and reward in time}
  \label{fig:coverage-reward}
\end{figure}
%% Create a subfigure of coverage-test, inside-test, inside-two-test and coverage-two-test

\begin{figure*}
  \centering
  \begin{subfigure}[b]{0.9\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-test.pdf}
    \caption{Coverage test}
    \label{fig:coverage-test}
  \end{subfigure}
  \begin{subfigure}[b]{0.9\linewidth}
    \includegraphics[width=\linewidth]{imgs/inside-test}
    \caption{Inside test}
    \label{fig:inside-test}
  \end{subfigure}
  \begin{subfigure}[b]{0.85\linewidth}
    \includegraphics[width=\linewidth]{imgs/inside-two-test}
    \caption{Inside two test}
    \label{fig:inside-two-test}
  \end{subfigure}
  \begin{subfigure}[b]{0.85\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-two-test}
    \caption{Coverage two test}
    \label{fig:coverage-two-test}
  \end{subfigure} 
  \caption{Test results, \ga{todo}}
  \label{fig:test}
\end{figure*}
\section{Conclusion and Future Work}
\ga{Page budget: 0.5}
\label{sec:conclusion}
